Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-07-15 doc/paper.tex
amta submission

% TODO: nodes class diagram? seems like an implementation detail...
% TODO: the fastest way i have to get dotted lines is through Word or PowerPoint, but I wanted to do the figures in OneNote, at least the camera-unready ones...
% TODO: it would be NICE if the tree matched a YAML example...
\begin{figure}[ht]
    \centering
    placeholder graphic, used to estimate space remaining in the paper
    \includegraphics{yaml-rb-rotated}

    \caption{
        An example tree made of nodes from the {\small \tt nodes} module.
        Thick lines indicate the ``constituent'' backbone that dictates the order of traversal in the {\small \tt generator} module, while thin lines indicate longer-range dependencies that are stored as links between nodes.
    }
    \label{fig:tree}

@@

The {\small \tt generator} module traverses the backbone of the ``interlingual'' tree from the {\small \tt nodes} module and recursively builds up parallel surface forms (bilingual sentence pairs).
Language-specific rules (e.g., word order) are implemented as ad hoc bits of code in this module,\footnote{
    Formally, the ``rule-based'' component of our system can be considered an extremely rudimentary synchronous (bilingual) context-free grammar of sorts \mycitep{transduction} that also admits long-range dependencies and semantic constraints.
    But it would be a disservice to those other well-developed methods to refer to our system as anything other than a makeshift way to generate multilingual sentences.
%    MEH: Also, example-based mode admits the possibility of just writing down translations and then generating parse trees on PORTIONS of those templates.
%    Although then I guess you have a transductive grammar on those portions?
} reading long-range dependency information from the tree as needed.
All language-dependent code is confined to this module, although this is coupled with language-dependent information from the YAML data (e.g., the tags and punctuation in Figure~\ref{fig:yaml-rb}). 
% TODO: More ideally, A desirable longer-term goal would be to move language rules to data, but that's beyond the scope of this paper.

@@

% maybe use a bulleted list for the different linguistic phenomena supported? yeah, let's just do that.

%Oh, maybe I should make this a table. that should cut down on the vertical spacing
%
\begin{table}[ht]
\centering
%\begin{tabular}{|l|l|} % tabularx allows automatic wrapping
\begin{tabularx}{\linewidth}{ |l|X| }
\hline
Clauses & indicative present active; intransitive, transitive, ``meta'' (clause subnode), ``modalish'' (e.g., ``Alice wants to kick Bob'' - with ``Alice'' linked to ``kick'' as in Figure~\ref{fig:tree} for subject-verb semantics) \\ \hline % see with a ``dependency link'' for subject-verb syntactic agreement and 
NP & noun phrases headed by nouns, names, or pronouns; optional modifiers as child nodes; pronoun antecedents via links \\ \hline
ADJP & adjectives as NP modifiers, with links to head nouns (e.g., this vs. these) \\ \hline
ADVP & adverbs as Clause or ADJP modifiers \\ \hline
Participles & transformations of clauses that convert subject child node into a linked node, allowing reuse of semantic constraints \\ \hline
PP & prepositional phrases (relations) between two nodes (noun/noun or noun/verb), with relation-dependent semantic requirements on both nodes; indirect objects as ``null preposition'' nodes \\ \hline
%Custom nodes & ``example-based'' nodes as discussed in Section~\ref{subsec:software} \\ \hline
%Transformations & alterations of existing templates - past tense, topicalization, \zh{把}-construction \\ \hline
\end{tabularx}
\caption{ 
    Linguistic constructs implemented for the ``rule-based'' mode of our generator.
}
\label{tab:implemented_linguistics}
\end{table}

%\begin{itemize}
%\item 
%\item participles as

%
%% other (if there's room)
%\item indirect object as a special case of ``null preposition'' - not necessarily linguistically well-founded, but good enough to make sentences

%\end{itemize}

@@

% SLIGHTLY inconsistent formatting, since now I have discussion ABOVE the table instead of below...
% but this creates a natural break for the discussion 
Table~\ref{tab:tst2010} shows the scores obtained by ``lightly'' post-editing the MERT tuning set {\small \tt tst2010} as follows. 
On the Chinese side, duplicated segments, applause lines, and translator notes were removed. 
Speaker annotations were normalized on both sides, as well as sentence alignments wherever possible.
As a result, small but noticeable differences in scores are observed. 
(For comparison, the difference between the best and worst scoring Chinese-to-English systems in IWSLT14 \mycitep{iwslt2014} was 6 BLEU points and 7 TER points.)
In this case, the post-edited tuning set seems to guide the MERT optimization procedure of Moses to a slightly different local optimum.

@@

\subsection{Data Snooping}
\textbf{
Warning: This section is dedicated to ``data snooping'' experiments that included information from the test set during training in one way or another.
The scores given here must not be considered indicative of actual out-of-sample system performance.
}

>>>>

% TODO: nodes class diagram? seems like an implementation detail...
% TODO: the fastest way i have to get dotted lines is through Word or PowerPoint, but I wanted to do the figures in OneNote, at least the camera-unready ones...
% TODO: preposition (or participle!) might be illustrative, but PP is just WAY too messy and involved to draw in a small space...
% TODO: it would be NICE if the tree matched a YAML example...
% TODO: oh, I could have made a tree diagram using some LaTeX nlp tree package? Word + pdf wound up being pretty fast though
\begin{figure}[ht]
    \centering
    \includegraphics{tree}
    \caption{
        An example tree made of nodes from the {\small \tt nodes} module.
        Solid lines indicate the ``constituent'' backbone that dictates the order of traversal in the {\small \tt generator} module, while dashed lines indicate longer-range syntactic or semantic dependencies that are stored as links between nodes.
        % In this example, V$_1$ can be a ``modalish'' verb like ``want'' that acts like a non-defective modal verb of sorts % meh, let's just bury it in the table below
    }
    \label{fig:tree}

@@

The {\small \tt generator} module traverses the backbone of the ``interlingual'' tree from the {\small \tt nodes} module and recursively builds up parallel surface forms (bilingual sentence pairs).
Language-specific rules (e.g., word order) are implemented as ad hoc bits of code in this module,\footnote{
    Formally, the ``rule-based'' component of our system can be considered an extremely rudimentary synchronous (bilingual) context-free grammar of sorts \mycitep{transduction} that also admits long-range dependencies and semantic constraints.
    But it would be a disservice to those other well-developed methods to refer to our system as anything other than a makeshift way to generate multilingual sentences.
%    MEH: Also, example-based mode admits the possibility of just writing down translations and then generating parse trees on PORTIONS of those templates.
%    Although then I guess you have a transductive grammar on those portions?
} reading long-range dependency information from the tree as needed.
All language-dependent code is confined to this module, although this is coupled with language-dependent information from the YAML data (e.g., the tags and punctuation in Figure~\ref{fig:yaml-rb}). 
For example, it is left to the generator to order prepositional phrases after their targets in English but before their targets in Chinese.
% TODO: More ideally, A desirable longer-term goal would be to move language rules to data, but that's beyond the scope of this paper.

@@

%Oh, maybe I should make this a table. that should cut down on the vertical spacing
%
\begin{table}[ht]
\centering
%\begin{tabular}{|l|l|} % tabularx allows automatic wrapping
\begin{tabularx}{\linewidth}{ |l|X| }
\hline
Clauses & indicative present active; intransitive, transitive, ``meta'' (clause subnode), ``modalish'' (e.g., ``Alice wants to kick Bob'' - with ``Alice'' linked to ``kick'' as in Figure~\ref{fig:tree} for subject-verb semantics) \\ \hline % see with a ``dependency link'' for subject-verb syntactic agreement and 
NP & noun phrases headed by nouns, names, or pronouns; optional modifiers as child nodes; pronoun antecedents via links \\ \hline
ADJP & adjectives as NP modifiers, with links to head nouns (e.g., this vs. these) \\ \hline
ADVP & adverbs as Clause or ADJP modifiers \\ \hline
VP & transformations of clauses into participles or modal(ish) complements; convert subject child node into a linked node, allowing reuse of semantic constraints \\ \hline
PP & prepositional phrases (relations) between two nodes (noun/noun or noun/verb), with relation-dependent semantic requirements on both nodes; indirect objects as ``null preposition'' nodes \\ \hline
%Custom nodes & ``example-based'' nodes as discussed in Section~\ref{subsec:software} \\ \hline
%Transformations & alterations of existing templates - past tense, topicalization, \zh{把}-construction \\ \hline
\end{tabularx}
\caption{ 
    Linguistic constructs implemented for the ``rule-based'' mode of our generator.
}
\label{tab:implemented_linguistics}
\end{table}

@@

% SLIGHTLY inconsistent formatting, since now I have discussion ABOVE the table instead of below...
% but this creates a natural break for the discussion 
Table~\ref{tab:tst2010} below shows the scores obtained by ``lightly'' post-editing the MERT tuning set {\small \tt tst2010} as follows. 
On the Chinese side, duplicated segments, applause lines, and translator notes were removed. 
Speaker annotations were normalized on both sides, as well as sentence alignments wherever possible.
As a result, small but noticeable differences in scores are observed. 
(For comparison, the difference between the best and worst scoring Chinese-to-English systems in IWSLT14 \mycitep{iwslt2014} was 6 BLEU points and 7 TER points.)
In this case, the post-edited tuning set seems to guide the MERT optimization procedure of Moses to a slightly different local optimum.

@@

\subsection{Data Snooping}
\textbf{Warning: This subsection is dedicated to ``data snooping'' experiments that included information from the test set during training in one way or another.
The scores given here must not be considered indicative of actual out-of-sample system performance.
}
